Pulling oral and IV data from CvTdb

### 2025-05-21 data-raw/pulling_oral_iv_data_cvtdb.Rmd
Updated data objects to reflect fixes to dose values for tamoxifen and MTBE data.

quick_stats(cvt_df)
quick_stats(cvt_2.0.0, old_version = TRUE)
# Note unique chemical counts

@@

# Dose Units: Need mg/kg... = ug/mg = ng/ug, and kg = L
# Conc Units: Need mg/L... = ug/mL = ng/uL = pg/nL, and kg = L, g = mL
cvt_df <- cvt_df %>% mutate(
  conc_units_normalized = stringr::str_replace(
    conc_units_original, "(?<=/m?)l", "L"
  ),
  loq_units = stringr::str_replace(loq_units, "(?<=/m?)l", "L"),
  invivPK_dose_level_units = "mg/kg",
  invivPK_conc_units = "ug/mL"
  )
# Set the equivalent units
equiv_mgkg <- c("mg/kg", "ug/g", "ng/mg")
equiv_mgL <- c("mg/L", "ug/mL", "ng/uL", "mg/kg", "mg/g", "ug/mg")

@@

@@

cvt_df %>% 
  filter(!(dose_level_units_original %in% equiv_mgkg),
                  dose_level_normalized == dose_level_original) %>%
  distinct(fk_study_id, 
           dose_level_original, dose_level_normalized,
           dose_level_units_original,
           administration_route_normalized, fk_extraction_document_id)

@@

cvt_df <- cvt_df %>% 
  mutate(
    invivPK_dose_level = case_when(
      # Standard weight of human (75kg)
      is.na(invivPK_dose_level) &
        is.na(weight_kg) &
        species %in% "human" &
        dose_level_units_original %in% "mg" ~ dose_level_original / 75,
      dose_level_units_original %in% "mg" ~ signif(
        invivPK_dose_level / weight_kg, digits = 5
      ),
      dose_level_units_original %in% "g" ~ signif(
        1E3 * invivPK_dose_level / weight_kg, digits = 5
      ),
      dose_level_units_original %in% "ug" ~ signif(
        1E-3 * invivPK_dose_level / weight_kg, digits = 5
      ),
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "ug/kg" ~ dose_level_original * 1E-3,
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "ng/kg" ~ dose_level_original * 1E-6,
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% c("mg/kg", "ug/mL") ~ dose_level_original,
      .default = invivPK_dose_level
    )
  )

@@

fit_pl <- plot(my_pk, best_fit = TRUE, log10_C = FALSE, use_scale_conc = TRUE)

fit_pl %>% filter(Chemical %in% "DTXSID1021116", Species %in% "human") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID9025617", Species %in% "mouse") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID9025617", Species %in% "rat") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID2021781", Species %in% "mouse") %>%
  pull(final_plot) %>% .[[1]]

>>>>

quick_stats(cvt_df)
# quick_stats(cvt_2.0.0, old_version = TRUE)
# Note unique chemical counts

@@

# Dose Units: Need mg/kg... = ug/mg = ng/ug, and kg = L
# Conc Units: Need mg/L... = ug/mL = ng/uL = pg/nL, and kg = L, g = mL

cvt_df <- cvt_df %>% mutate(
  conc_units_normalized = stringr::str_replace(
    conc_units_original, "(?<=/m?)l", "L"
  ),
  loq_units = stringr::str_replace(loq_units, "(?<=/m?)l", "L"),
  invivPK_dose_level_units = "mg/kg",
  invivPK_conc_units = "ug/mL"
  )
# Set the equivalent units
equiv_mgkg <- c("mg/kg", "ug/g", "ng/mg")
equiv_mgL <- c("mg/L", "ug/mL", "ng/uL", "mg/kg", "mg/g", "ug/mg")

@@

cvt_df %>% filter(stringr::str_detect(dose_level_units_original, "L")) %>%
  distinct(analyzed_chem_dtxsid, analyzed_chem_name_original,
           dose_level_original, dose_level_normalized,
           dose_level_units_original,
           dose_volume, dose_volume_units,
           administration_route_normalized,
           pmid, fk_extraction_document_id,
           other_study_identifier) %>%
  glimpse()

@@

@@

cvt_df <- cvt_df %>% 
  mutate(
    invivPK_dose_level = case_when(
      # Standard weight of human (75kg)
      is.na(invivPK_dose_level) &
        is.na(weight_kg) &
        species %in% "human" &
        dose_level_units_original %in% "mg" ~ dose_level_original / 75,
      dose_level_units_original %in% "mg" ~ signif(
        invivPK_dose_level / weight_kg, digits = 5
      ),
      dose_level_units_original %in% "g" ~ signif(
        1E3 * invivPK_dose_level / weight_kg, digits = 5
      ),
      dose_level_units_original %in% "ug" ~ signif(
        1E-3 * invivPK_dose_level / weight_kg, digits = 5
      ),
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "ug/kg" ~ dose_level_original * 1E-3,
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "ng/kg" ~ dose_level_original * 1E-6,
      # There are some cases where there is a volume given (not mass normalized)
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "ug/mL" &
        !is.na(dose_volume) & 
        dose_volume_units %in% "mL/kg" ~ dose_level_original * dose_volume / 1E3,
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "ug/mL" &
        !is.na(dose_volume) & 
        dose_volume_units %in% "mL" &
        species %in% "human" ~ dose_level_original * dose_volume / (75E3),
      is.na(invivPK_dose_level) &
        dose_level_units_original %in% "mg/kg" ~ dose_level_original,
      .default = invivPK_dose_level
    )
  )

@@

fit_pl <- plot(my_pk, log10_C = FALSE, use_scale_conc = TRUE)

fit_pl %>% filter(Chemical %in% "DTXSID1021116", Species %in% "human") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID9025617", Species %in% "mouse") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID9025617", Species %in% "rat") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID2021781", Species %in% "mouse") %>%
  pull(final_plot) %>% .[[1]]
fit_pl %>% filter(Chemical %in% "DTXSID1034187") %>% # Tamoxifen sanity check
  pull(final_plot) %>% .[[1]]

get_winning_model(my_pk) %>% filter(Chemical %in% "DTXSID1034187")
